Effect of quality-adjusted production function inputs

### 2015-04-24 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Can now use saved model data.

#####################################################
evalCES=TRUE
#####################################################
@

@@

<<Formulas>>=
# Set up the formulas for each type of analysis.
f_PT_unadj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_unadj, 
                    l = l_PT_unadj, t = t_PT)))
f_PT_unadj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_unadj,
                    l = l_PT_unadj, e = e_PT_unadj, t = t_PT)))
f_PT_adj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, t = t_PT)))
f_PT_adj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, e = e_PT_adj, t = t_PT)))
f_UK_unadj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, t = t_UK)))
f_UK_unadj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, e = e_UK_unadj, t = t_UK)))
f_UK_adj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, t = t_UK)))
f_UK_adj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, e = e_UK_adj, t = t_UK)))
@

@@

<<Portugal C-D undadjusted without energy>>=
ptCDunadjNoE <- cdModel(formula = f_PT_unadj_noE, 
                        data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
@

<<Portugal C-D undadjusted with energy>>=
ptCDunadjWithE <- cdModel(formula = f_PT_unadj_withE, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjWithE)
@

The Solow residual (\texttt{lambda}) is always negative. 
However, the absolute value of \texttt{lambda} decreases from 
\Sexpr{naturalCoef(ptCDunadjNoE)$lambda} to 
\Sexpr{naturalCoef(ptCDunadjWithE)$lambda}
when adding energy.
So it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the absolute value of the Solow residual somewhat.

@@

<<Portugal C-D adjusted without energy>>=
ptCDadjNoE <- cdModel(formula = f_PT_adj_noE, 
                      data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
@

<<Portugal C-D adjusted with energy>>=
ptCDadjWithE <- cdModel(formula = f_PT_adj_withE, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjWithE)
@

Using quality-adjusted inputs, the Solow residual (\texttt{lambda}) moves from 
\Sexpr{naturalCoef(ptCDadjNoE)$lambda} to
\Sexpr{naturalCoef(ptCDadjWithE)$lambda} when including energy.
The absolute value of the Solow residual again decreases when adding energy.
But, with quality-adjusted inputs, the effect is much stronger.

@@

\begin{table} \caption{Portugal Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCDunadjNoE)$lambda}      & \Sexpr{naturalCoef(ptCDadjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCDunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCDadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK C-D>>=
ukCDunadjNoE <- cdModel(formula = f_UK_unadj_noE, 
                        data = dataUK, constrained = TRUE)
ukCDunadjWithE <- cdModel(formula = f_UK_unadj_withE, 
                          data = dataUK, constrained = TRUE)
ukCDadjNoE <- cdModel(formula = f_UK_adj_noE, 
                        data = dataUK, constrained = TRUE)
ukCDadjWithE <- cdModel(formula = f_UK_adj_withE, 
                          data = dataUK, constrained = TRUE)
@

\begin{table} \caption{UK Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ukCDunadjNoE)$lambda}      & \Sexpr{naturalCoef(ukCDadjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ukCDunadjWithE)$lambda}    & \Sexpr{naturalCoef(ukCDadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(ukCDadjWithE)$lambda * 100}%,
a very small number indeed.
For the UK, it appears that both quality-adjustment and addition of energy as a factor of production
can effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(ukCDadjWithE)$lambda}.

Interestingly, the Solow residual acutally increases from 
\Sexpr{naturalCoef(ukCDunadjNoE)$lambda} to
\Sexpr{naturalCoef(ukCDunadjWithE)$lambda}
for unadjusted factors of production when energy is added.
To find out why, let's look at all the coefficients.

<<UK CD Unadjusted Coefficients>>=
naturalCoef(ukCDunadjNoE)
naturalCoef(ukCDunadjWithE)
@

When energy is included, much of the effect of capital (\texttt{alpha\_1}) is 
attributed to energy (\texttt{alpha\_3}), 
as \texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(ukCDunadjNoE)$alpha_1} to
\Sexpr{naturalCoef(ukCDunadjWithE)$alpha_1}.

@@

<<UK CD Quality-adjusted Coefficients>>=
naturalCoef(ukCDadjNoE)
naturalCoef(ukCDadjWithE)
@

\texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(ukCDadjNoE)$alpha_1} to
\Sexpr{naturalCoef(ukCDadjWithE)$alpha_1}.

@@

<<PT CES-kl, warning=FALSE, eval=evalCES>>=
ptCESklunadjNoE <- cesModel(formula = f_PT_unadj_noE, 
                            nest = c(1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESklunadjNoE)
ptCESkladjNoE <- cesModel(formula = f_PT_adj_noE, 
                          nest = c(1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkladjNoE)
@

@@

First, look at unadjusted inputs using the ($kl$)$e$ nesting.

<<PT CES-kle unadjusted, warning=FALSE, eval=evalCES>>=
ptCESkleunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleunadjWithE)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESkleadjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESkleunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESkleadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

The Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).

@@

<<PT CES-lek unadjusted, warning=FALSE, eval=evalCES>>=
ptCESlekunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekunadjWithE)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESlekadjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESlekunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESlekadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

The Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).

%------------------------------
\subsubsection{($ek$)$k$ nesting} 
%------------------------------

@@

<<PT CES-ekl unadjusted, warning=FALSE, eval=evalCES>>=
ptCESeklunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESeklunadjWithE)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESekladjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESekladjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESeklunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESekladjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

The Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).

>>>>

#####################################################
rerunModels=FALSE
#####################################################
@

@@

<<Formulas>>=
# Set up the formulas for each type of analysis.
f <- NULL
f$PT$unadj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_unadj, 
                    l = l_PT_unadj, t = t_PT)))
f$PT$unadj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_unadj,
                    l = l_PT_unadj, e = e_PT_unadj, t = t_PT)))
f$PT$adj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, t = t_PT)))
f$PT$adj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, e = e_PT_adj, t = t_PT)))
f$UK$unadj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, t = t_UK)))
f$UK$unadj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, e = e_UK_unadj, t = t_UK)))
f$UK$adj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, t = t_UK)))
f$UK$adj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, e = e_UK_adj, t = t_UK)))
@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  nest <- NULL
  nest$kl <- c(1,2)
  nest$kle <- c(1,2,3)
  nest$lek <- c(2,3,1)
  nest$ekl <- c(3,1,2)
  models <- NULL
  # Portugal, CD
  models$PT$unadj$noE$CD <- cdModel(formula = f$PT$unadj$noE, 
                                    data = dataPT, constrained = TRUE)
  models$PT$unadj$withE$CD <- cdModel(formula = f$PT$unadj$withE, 
                                      data = dataPT, constrained = TRUE)
  models$PT$adj$noE$CD <- cdModel(formula = f$PT$adj$noE, 
                        data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CD <- cdModel(formula = f$PT$adj$withE, 
                          data = dataPT, constrained = TRUE)
  # UK, CD
  models$UK$unadj$noE$CD <- cdModel(formula = f$UK$unadj$noE, 
                          data = dataUK, constrained = TRUE)
  models$UK$unadj$withE$CD <- cdModel(formula = f$UK$unadj$withE, 
                            data = dataUK, constrained = TRUE)
  models$UK$adj$noE$CD <- cdModel(formula = f$UK$adj$noE, 
                        data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CD <- cdModel(formula = f$UK$adj$withE, 
                          data = dataUK, constrained = TRUE)
  # PT, CESkl
  models$PT$unadj$noE$CESkl <- cesModel(formula = f$PT$unadj$noE, 
                              nest = nest$kl, data = dataPT, constrained = TRUE)
  models$PT$adj$noE$CESkl <- cesModel(formula = f$PT$adj$noE, 
                            nest = nest$kl, data = dataPT, constrained = TRUE)
  # PT, CESkle
  models$PT$unadj$withE$CESkle <- cesModel(formula = f$PT$unadj$withE, 
                                 nest = nest$kle, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESkle <- cesModel(formula = f$PT$adj$withE, 
                               nest = nest$kle, data = dataPT, constrained = TRUE)
  # PT, CESlek
  models$PT$unadj$withE$CESlek <- cesModel(formula = f$PT$unadj$withE, 
                                 nest = nest$lek, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESlek <- cesModel(formula = f$PT$adj$withE, 
                               nest = nest$lek, data = dataPT, constrained = TRUE)
  # PT, CESekl
  models$PT$unadj$withE$CESekl <- cesModel(formula = f$PT$unadj$withE, 
                                 nest = nest$ekl, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESekl <- cesModel(formula = f$PT$adj$withE, 
                               nest = nest$ekl, data = dataPT, constrained = TRUE)
  # UK, CESkl
  models$UK$unadj$noE$CESkl <- cesModel(formula = f$UK$unadj$noE, 
                              nest = nest$kl, data = dataUK, constrained = TRUE)
  models$UK$adj$noE$CESkl <- cesModel(formula = f$UK$adj$noE, 
                            nest = nest$kl, data = dataUK, constrained = TRUE)
  # UK, CESkle
  models$UK$unadj$withE$CESkle <- cesModel(formula = f$UK$unadj$withE, 
                                 nest = nest$kle, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESkle <- cesModel(formula = f$UK$adj$withE, 
                               nest = nest$kle, data = dataUK, constrained = TRUE)
  # UK, CESlek
  models$UK$unadj$withE$CESlek <- cesModel(formula = f$UK$unadj$withE, 
                                 nest = nest$lek, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESlek <- cesModel(formula = f$UK$adj$withE, 
                               nest = nest$lek, data = dataUK, constrained = TRUE)
  # UK, CESekl
  models$UK$unadj$withE$CESekl <- cesModel(formula = f$UK$unadj$withE, 
                                 nest = nest$ekl, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESekl <- cesModel(formula = f$UK$adj$withE, 
                               nest = nest$ekl, data = dataUK, constrained = TRUE)
  # Save models to disk.
  save(models, file=fileName)
}

@

@@

<<Portugal C-D undadjusted without energy>>=
naturalCoef(models$PT$unadj$noE$CD)
@

<<Portugal C-D undadjusted with energy>>=
naturalCoef(models$PT$unadj$withE$CD)
@

The Solow residual (\texttt{lambda}) is always negative. 
However, the absolute value of \texttt{lambda} decreases from 
\Sexpr{naturalCoef(models$PT$unadj$noE$CD)$lambda} to 
\Sexpr{naturalCoef(models$PT$unadj$withE$CD)$lambda}
when adding energy.
So it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the absolute value of the Solow residual somewhat.

@@

<<Portugal C-D adjusted without energy>>=
naturalCoef(models$PT$adj$noE$CD)
@

<<Portugal C-D adjusted with energy>>=
naturalCoef(models$PT$adj$withE$CD)
@

Using quality-adjusted inputs, the Solow residual (\texttt{lambda}) moves from 
\Sexpr{naturalCoef(models$PT$adj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CD)$lambda} when including energy.
The absolute value of the Solow residual again decreases when adding energy.
But, with quality-adjusted inputs, the effect is much stronger.

@@

\begin{table} \caption{Portugal Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CD)$lambda}    & \Sexpr{naturalCoef(models$PT$adj$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CD)$lambda}  & \Sexpr{naturalCoef(models$PT$adj$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK C-D>>=
@

\begin{table} \caption{UK Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda}   & \Sexpr{naturalCoef(models$UK$adj$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda} & \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda * 100}%,
a very small number indeed.
For the UK, it appears that both quality-adjustment and addition of energy as a factor of production
can effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda}.

Interestingly, the Solow residual acutally increases from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda}
for unadjusted factors of production when energy is added.
To find out why, let's look at all the coefficients.

<<UK CD Unadjusted Coefficients>>=
naturalCoef(models$UK$unadj$noE$CD)
naturalCoef(models$UK$unadj$withE$CD)
@

When energy is included, much of the effect of capital (\texttt{alpha\_1}) is 
attributed to energy (\texttt{alpha\_3}), 
as \texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$alpha_1}.

@@

<<UK CD Quality-adjusted Coefficients>>=
naturalCoef(models$UK$adj$noE$CD)
naturalCoef(models$UK$adj$withE$CD)
@

\texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$adj$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$alpha_1}.

@@

<<PT CES-kl, warning=FALSE>>=
naturalCoef(models$PT$unadj$noE$CESkl)
naturalCoef(models$PT$adj$noE$CESkl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals,
\Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the Solow residual.

@@

Now, look at unadjusted inputs using the ($kl$)$e$ nesting.

<<PT CES-kle unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESkle)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESkle)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESkle)$lambda} 
                                                                    & \Sexpr{naturalCoef(models$PT$adj$withE$CESkle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

In contrast to the Cobb-Douglas model,
the CES model with ($kl$)$e$ nesting
produces positive Solow residuals in all cases. 
We also see that both 
quality-adjusting the inputs and
including energy
reduce the Solow residual (\texttt{lambda}).

@@

<<PT CES-lek unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESlek)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESlek)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESlek)$lambda} & \Sexpr{naturalCoef(models$PT$adj$withE$CESle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
After energy has been added, quality-adjustment 
has very little effect on the Solow residual,
creating a slight increase from 
\Sexpr{naturalCoef(models$PT$unadj$withE$CESlek)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CESlek)$lambda}.

%------------------------------
\subsubsection{($ek$)$l$ nesting} 
%------------------------------

@@

<<PT CES-ekl unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESekl)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESekl)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESekl)$lambda} 
                                                                          & \Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the ($ek$)$l$ nesting, 
the Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).
Adding energy decreases the Solow residual 
(from \Sexpr{naturalCoef(models$PT$unadj$withE$CESekl)$lambda} to 
\Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda})
when enegy is included in the model.
But, adding energy slightly increases the Solow residual
(from \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} to \Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda})
when quality-adjusted inputs are used.

%++++++++++++++++++++++++++++++
\subsection{Do we obtain a positive Solow residual for the UK with the CES model?} 
\label{sec:Solow_positive_for_UK_in_CES?}
%++++++++++++++++++++++++++++++

<<UK CES-kl, warning=FALSE>>=
naturalCoef(models$UK$unadj$noE$CESkl)
naturalCoef(models$UK$adj$noE$CESkl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
\Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.

Now, look at unadjusted inputs using the ($kl$)$e$ nesting.

<<UK CES-kle unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESkle)
@

<<UK CES-kle quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESkle)
@

Table~\ref{tab:lambda_CES-kle_UK} summarizes the UK Solow residual results with CES and the ($kl$)$e$ nesting.

\begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESkle)$lambda} 
                                                                & \Sexpr{naturalCoef(models$UK$adj$withE$CESkle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Like the Cobb-Douglas model, the CES model produces positive Solow residuals
with the ($kl$)$e$ nesting.
We also see that both 
quality-adjusting the inputs and
including energy
reduce the Solow residual (\texttt{lambda}).

<<UK CES-lek unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESlek)
@

<<UK CES-lek quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESlek)
@

Table~\ref{tab:lambda_CES-lek_UK} summarizes the UK Solow residual results with CES and the ($le$)$k$ nesting.

\begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESlek)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CESlek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
Both quality-adjustment and inclusion of energy 
decrease the magnitude of the Solow residual.

%------------------------------
\subsubsection{($ek$)$l$ nesting}
%------------------------------

Next, look at unadjusted inputs using the ($el$)$l$ nesting.

<<UK CES-ekl unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESekl)
@

<<UK CES-ekl quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESekl)
@

Table~\ref{tab:lambda_CES-ekl_UK} summarizes the UK Solow residual results with CES and the ($ek$)$l$ nesting.

\begin{table} \caption{UK Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESekl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CESekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($ek$)$l$ nesting,
the CES model with ($ek$)$l$ nesting
produces positive Solow residuals in all cases. 
Both quality-adjustment and inclusion of energy 
decrease the magnitude of the Solow residual.
